---
title: "Code to replicate analysis of: Tools of the Trade – When Are Software Tools Mentioned in Computational Text Analysis Research?"
format:
  html:
    embed-resources: true
knitr:
  opts_chunk: 
    collapse: true
keywords:
  - Computational Text Analysis
  - Research Software 
  - Open Science
  - Software Citation
---

# Preliminiaries: loading data and libraries
```{r}
#| output: false
library(arrow)
library(dplyr)
library(tidyr)
library(haven)
library(knitr)
library(scales)
library(xtable)
```
And here are two helper functions for later: 
The first one checks whether the programming languages column contains a language one is looking for (e.g. is this written in Python?).
The second combines our different measures of accessbility into an additive index.
```{r}
#| output: false
check_programming_language <- function(column_of_interest, language_of_interest){
  if (is.null(language_of_interest)){
    if (length(unlist(column_of_interest)) == 0){
      return(TRUE)
    }
    else {
      return(FALSE)
    }
  }
  for (language in unlist(column_of_interest)){
    if (language == language_of_interest){
      return(TRUE)
    }
  }
  return(FALSE)
}

build_accessible_index <- function(input_data){
  value = 0
  if (input_data$dummy_doi){
    value = value+1
  }
  if (input_data$dummy_materials){
    value = value+1
  }
  if((input_data$dummy_cran) | (input_data$dummy_pypi)|  (input_data$dummy_github)){
    value = value+1
  }
  if (input_data$graphical_user_interface){
    value = value+1
  }
  if (input_data$user_access == "free"){
    value = value+1
  }
  return(value)
}
```
```{r}
articles <- read_feather("articles.feather")

tools <- read_feather("tools.feather") %>% 
  rowwise() %>%
  mutate(
    dummy_python = check_programming_language(programming_languages, 
                                              language_of_interest = "python"),
    dummy_r = check_programming_language(programming_languages, 
                                         language_of_interest = "r"),
    dummy_nolang = check_programming_language(programming_languages, 
                                              language_of_interest = NULL),
    # some github links are spelled as "quanteda/quanteda/" in Meteor
    github = sub("\\/$", "", github)) %>%
  mutate(
    open_source = case_when(
      open_source == "yes" ~ TRUE,
      open_source == "no" ~ FALSE,
      TRUE ~ FALSE
    ),
    dummy_doi = case_when(
      is.na(doi) ~ FALSE,
      !is.na(doi) ~ TRUE,
      TRUE ~ FALSE
    ),
    dummy_github = case_when(
      is.na(github) ~ FALSE,
      !is.na(github) ~ TRUE,
      TRUE ~ FALSE
    ),
    dummy_cran = case_when(
      is.na(cran) ~ FALSE,
      !is.na(cran) ~ TRUE,
      TRUE ~ FALSE
    ),
    dummy_pypi = case_when(
      is.na(pypi) ~ FALSE,
      !is.na(pypi) ~ TRUE,
      TRUE ~ FALSE
    ),
    dummy_materials = case_when(
      !is.null(materials) ~ TRUE,
      TRUE ~ FALSE
    ),
   index_accessible = build_accessible_index(.data)
  )
```

This is the shape of the article data: 
```{r}
kable(head(articles))
```

and of the tool data (without kable, as the fields are too wide): 
```{r}
head(tools)
```

# Descriptive Tables 

Proportion of articles per year and the reporting of software tools
```{r}
articles %>%
  group_by(year) %>%
  mutate(articles_per_year = n()) %>%
  separate_rows(tool_ids, sep=",\\s*") %>%
  reframe("Publications" = articles_per_year,
          "Percentage Naming Tools"  = (articles_per_year - sum(is.na(tool_ids)))/articles_per_year*100,
            "Tools named" = n(),
            "Tools per publication" = n()/articles_per_year,
            "Tools per publication which has tools" = n()/(articles_per_year - sum(is.na(tool_ids)))
          ) %>%
  distinct() %>% 
  kable()
```

Comparison of journal reporting trends for the top communication science journals
```{r}
articles %>%
  # exchange this filter to have all journals, or to focus on any other discipline
  filter(top20_soc == 1) %>%
  group_by(journal) %>%
  mutate(n_per_journals = n()) %>% 
  filter(n_per_journals > 4)  %>%
  reframe("Number of Articles" = n_per_journals, 
            "Percentage Naming Tools"  =round(((n_per_journals - sum(is.na(tool_ids)))/n_per_journals), digits = 2),
            "Percentage Reporting/Discussing Validation" = sum(validation_any_boolean, na.rm = TRUE)/n_per_journals) %>% 
  arrange(desc(`Percentage Naming Tools`)) %>% 
  distinct() %>% 
  mutate(
    "journal" = haven::as_factor(journal),
    "Percentage Naming Tools" = percent(`Percentage Naming Tools`),
    "Percentage Reporting/Discussing Validation" = percent(`Percentage Reporting/Discussing Validation`)
  ) %>% 
  select(journal, "Percentage Naming Tools", "Percentage Reporting/Discussing Validation", "Number of Articles")  %>%
  xtable()
```

Most frequently reported software tools
```{r}
tools %>% 
  arrange(desc(n)) %>% 
  reframe("Software Tool Name" = name, 
        "Number of mentions" = n) %>% 
  head() %>% 
  kable()
```

Distribution of accessibility of software tools

```{r}
tools %>%
  group_by(index_accessible) %>%
  summarise(Percentage = n() / nrow(tools) * 100,
  Mean = mean(n),
  SD = sd(n)) %>% 
  kable()
```

# Hypothesis Testing 

Create a dichotomous variable whether an article mentions tools or not, and combine both codings of validation into one.
```{r}
articles <- articles %>% 
  mutate(tools_boolean = case_when(
      number_tools > 0 ~ TRUE,
      number_tools == 0 ~ FALSE,
    TRUE ~ FALSE),
    validation_disc_boolean = case_when(
      valid1 > 0 ~ TRUE,
      valid1 == 0 ~ FALSE,
      TRUE ~ FALSE),
    validation_data_boolean = case_when(
      valid2 > 0 ~ TRUE,
      valid2 == 0 ~ FALSE,
      TRUE ~ FALSE),
    validation_any_boolean = case_when(
      validation_disc_boolean | validation_data_boolean ~ TRUE, 
      TRUE ~ FALSE)
    )
```

*H1 The more recent a journal article was published, the more likely it is to mention software tools used.*

```{r}
chisq.test(articles$tools_boolean, articles$year)
```

*H2a: There are differences between disciplines regarding whether software tools are mentioned.*

```{r}
chisq.test(articles$tools_boolean, articles$top20_com)
chisq.test(articles$tools_boolean, articles$top20_pol)
chisq.test(articles$tools_boolean, articles$top20_psychplus)
chisq.test(articles$tools_boolean, articles$top20_socplus)
```

*H2b: There are differences between journals within the same discipline regarding whether software tools are mentioned.*

```{r}
disciplines = c("top20_com", "top20_pol", "top20_socplus", 
                "top20_psychplus", "top20_soc")
for (discipline in disciplines){
  print(discipline)
  only_one_discipline_articles <- articles %>%  
    dplyr::filter(haven::as_factor(.data[[discipline]]) == "yes")
  print(chisq.test(only_one_discipline_articles$tools_boolean, 
                  only_one_discipline_articles$journal))
}
```

*H3: Journal articles that discuss or demonstrate the validity of their methodology are more likely to report the associated software tools.*

```{r}
chisq.test(articles$tools_boolean, articles$validation_any_boolean)
```

*Appendix for H3: Journals that have a higher share of articles discussing validity of their methodology have a higher average share of articles mentioneing their associated software tools.*

```{r}
articles_corelation_h4 <- articles %>%
  group_by(journal) %>%
  mutate(no_articles = n()) %>% 
  mutate("avg_naming_tools"  =((no_articles - sum(is.na(tool_ids)))/no_articles),
            "avg_naming_valid" = sum(validation_any_boolean, na.rm = TRUE)/no_articles) %>% 
  select(journal, no_articles, avg_naming_tools, avg_naming_valid) %>% 
  distinct() 
cor.test(articles_corelation_h4$avg_naming_tools, articles_corelation_h4$avg_naming_valid, method = "kendall")
```

*H4: Increased software tool accessibility increases the frequency of its mentions in journal articles.*

```{r}
cor.test(tools$index_accessible, tools$n, method = "kendall")
```

*H5: There is a positive correlation between the development activity of a software tool and the number of journal article mentions of a software tool.*

```{r}
cor.test(tools$issues_relation_closed_open, tools$n, method = "kendall")
cor.test(tools$mean_commits, tools$n, method = "kendall")
```
